---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# p2distance

<!-- badges: start -->
[![CRAN status](https://www.r-pkg.org/badges/version/p2distance)](https://CRAN.R-project.org/package=p2distance)
[![pkgdown](https://github.com/ajpelu/p2distance/actions/workflows/pkgdown.yaml/badge.svg)](https://github.com/ajpelu/p2distance/actions/workflows/pkgdown.yaml)
<!-- badges: end -->

`p2distance` computes the **P2 distance** synthetic indicator (Peña, 1977), a
method for combining several partial indicators — quality of life, welfare,
environmental quality, development, and so on — into a single measure that
lets you compare different entities (countries, regions, cities...) on a
common scale.

Unlike Principal Component Analysis or other traditional aggregation
methods, P2 distance avoids arbitrary weighting: each variable is
weighted by how much *new* information it contributes once the variables
already included are accounted for, using an iterative correction based on
each variable's coefficient of determination (R²).

## Installation

Install the released version from CRAN:

``` r
install.packages("p2distance")
```

Or the development version from GitHub:

``` r
# install.packages("pak")
pak::pak("ajpelu/p2distance")
```

## Example

`p2distance` ships with `welfare`, a dataset of 20 quality-of-life
indicators for the 27 countries of the European Union (2002-2007,
EurLIFE). Let's rank countries by an overall welfare indicator:

```{r example}
library(p2distance)

data(welfare)
welfare_mat <- as.matrix(welfare)

ind <- p2distance(welfare_mat, reference_vector_function = min, iterations = 20)

# Ranking (higher P2 distance = further from the reference = lower welfare)
sort(ind$p2distance[, 1])
```

Each variable's contribution to the indicator can also be inspected —
here, how strongly each partial indicator correlates with the overall
P2 distance:

```{r cor-plot, fig.alt = "Barplot of correlations between each welfare variable and the P2 distance indicator"}
barplot(
  sort(ind$cor.coeff[, 1]),
  horiz = TRUE, las = 1, col = "steelblue", border = NA,
  xlab = "Correlation with P2 distance indicator",
  cex.names = 0.7
)
```

```{r p2-plot, fig.alt = "Barplot of P2 distance indicator by country, sorted"}
barplot(
  sort(ind$p2distance[, 1]),
  las = 3, cex.names = 0.7, col = "white",
  ylab = "P2 distance (welfare indicator)"
)
```

## Background

The P2 distance was proposed by the Spanish economist Jesús B. Peña Trapero
to measure social welfare (Peña, 1977), and has since been applied to
environmental quality indices, regional development, and inequality
studies (see `?p2distance` for the full formula and additional references).

## References

Peña, J. B. (1977). *Problemas de la medición del bienestar y conceptos
afines (una aplicación al caso Español)*. Madrid: Instituto Nacional de
Estadística (INE).

Peña, J. B. (2009). La medición del bienestar social: una revisión
crítica. *Estudios de Economía Aplicada*, 27(2), 299–324.

## Citation

If you use this package, please cite it — see `citation("p2distance")`.

## Code of Conduct

Please note that the p2distance project is released with a
[Contributor Code of Conduct](https://contributor-covenant.org/version/2/1/CODE_OF_CONDUCT.html).
By contributing to this project, you agree to abide by its terms.