---
title: "Impact"
bibliography: citing-works.bib
csl: apa.csl
nocite: '@*'
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(p2distance)
```


## Downloads

Since being restored to CRAN, `p2distance` downloads are tracked via the
`cranlogs` package. The plot below shows cumulative downloads over time

```{r downloads, message = FALSE, warning = FALSE, echo=FALSE, fig.width=10, fig.height=7}
library(cranlogs)
library(dplyr)
library(tidyr)
library(lubridate)
library(ggplot2)

p2d <- cranlogs::cran_downloads("p2distance", from = "2010-01-01")

p2d_stats <- p2d |>
  tidyr::drop_na() |>
  filter(count > 0)

monthly <- p2d_stats |>
  mutate(month = lubridate::ceiling_date(date, "month") - 1) |>
  group_by(month) |>
  summarise(downloads = sum(count), .groups = "drop") |>
  mutate(csum = cumsum(downloads))

monthly |>
  ggplot(aes(x = month, y = csum)) +
  geom_line(color = "steelblue", linewidth = 1) + 
  scale_x_date(date_breaks = "1 year", date_labels = "%Y", limits = c(as.Date("2012-01-01"), as.Date("2026-01-01"))) +
  labs(x = NULL, y = "Cumulative downloads",
       title = "Cumulative CRAN downloads of p2distance") +
  theme_minimal() +
  theme(panel.grid.minor = element_blank())
```

## Citing works 

The following peer-reviewed studies have used `p2distance` to compute
synthetic well-being, competitiveness, or sustainability indicators, listed in 
chronological order.

Is your work missing from this list? [Open an issue](https://github.com/ajpelu/p2distance/issues)
or send a pull request adding it.
